---
title: "P8105 Homework 5"
author: "Junxian Chen (jc5314)"
date: "11/2/2019"
output: github_document
---

```{r, include = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  message = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
theme_set(theme_gray(base_size = 10) + theme(legend.position = "bottom"))
```

# Problem 1

```{r}
set.seed(10)

iris_with_missing = iris %>% 
  map_df(~replace(.x, sample(1:150, 20), NA)) %>%
  mutate(Species = as.character(Species))
```

Firstly, load the dataset with missing values.

```{r}
replaceNA = function(x){
  
  if (class(x) == "numeric") {
    replace_na(x, mean(x, na.rm = TRUE))
  } else if (class(x) == "character") {
    replace_na(x, "virginica") 
  } else {
    stop("Input is neither numeric nor character")
  }

}
```

Next, a function called `replaceNA` is created. This function will take a vector as an argument, detect the input variable type and then replace missing values using the rules defined in the question. If the input value is  neither numeric nor character, the function will stop and give out error information.

Lastly, apply this function to the columns of dataset using a map statement.

```{r}
output = map_df(iris_with_missing, replaceNA)

output
```


# Problem 2

Read in and tidy the datasets:

```{r}
p2_df =
  tibble(
    file_name = list.files(path = './data', full.names = TRUE),
    data = map(file_name, ~read_csv(.)),
    group = case_when(
      file_name %>% str_detect("con") ~ "control",
      file_name %>% str_detect("exp") ~ "experiment",
    ),
    id = as.numeric(str_extract_all(file_name, "[0-9]+"))
  ) %>% 
  unnest(cols = c(data)) %>% 
  pivot_longer(
    week_1:week_8,
    names_to = "week",
    values_to = "data"
  ) %>% 
  mutate(
    full_id = paste(group, "_", as.character(id), sep = ""),
    week = as.numeric(str_extract_all(week, "[0-9]+"))
  ) %>% 
  select(full_id, group, id, week, data)

p2_df
```

A spaghetti plot showing observations on each subject over time:

```{r}
p2_df %>% 
  ggplot(aes(x = week, y = data, group = full_id, color = group)) +
  geom_line() +
  xlab("Week") +
  ylab("Data")
```

*Comments:* The data values of the experiment group were increasing along with time and the overall data values in experiment group were distinctly larger than the control group after 6 weeks.

# Problem 3

```{r}
generate_xy = function(beta_1, n, beta_0, sigma){
  sim_data = tibble(
    x = rnorm(n),
    y = beta_0 + beta_1 * x + rnorm(n, 0, sigma)
  )
  return(sim_data)
}
```

```{r}
generate_df = function(beta_1, generate_xy, n = 30, beta_0 = 2, sigma = sqrt(50)){
  
  tibble(
    truth = beta_1,
    data = rerun(10, generate_xy(beta_1, n, beta_0, sigma)),
    model = map(data, ~ broom::tidy(lm(y ~ x, data = .))),
    estimate = map(model, ~ .x[2,2]),
    p_value = map(model, ~ .x[2,5])
  )  %>% 
  unnest(estimate:p_value) %>% 
  select(-model, p_value = p.value)
  
}
```


```{r}
set.seed(1)

df_1 = generate_df(beta_1 = 0, generate_xy)
```

```{r}
beta_1 = list(1, 2, 3, 4, 5, 6)

df_2 = 
  map(beta_1, ~ generate_df(.x, generate_xy)) %>% 
  bind_rows()
```

* A plot showing the proportion of times the null was rejected (the power of the test) on the $y$ axis and the true value of $\beta_1$ on the $x$ axis.

```{r}
df_2 %>% 
  group_by(truth) %>% 
  mutate(
    reject_prop = sum(p_value < 0.05) / length(p_value)
  ) %>% 
  ggplot(aes(x = truth, y = reject_prop, fill = truth)) +
  geom_col() + 
  xlab("True Value of Beta_1") +
  ylab("Proportion of times the null was rejected") +
  scale_x_continuous(
    breaks = c(1, 2, 3, 4, 5, 6)
  )
```

*Comments: Describe the association between effect size and power.*

* Make a plot showing the average estimate of $\hat{\beta_1}$ on the $y$ axis and the true value of $\beta_1$ on the $x$ axis.

```{r}
df_2 %>% 
  group_by(truth) %>% 
  summarize(mean_estimator = mean(estimate)) %>% 
  ggplot(aes(x = truth, y = mean_estimator, fill = truth)) +
  geom_col() + 
  xlab("True Value of Beta_1") +
  ylab("Average estimate of Beta_1") +
  scale_x_continuous(
    breaks = c(1, 2, 3, 4, 5, 6)
  )
```


```{r}
df_2 %>% 
  filter(p_value < 0.05) %>% 
  group_by(truth) %>% 
  summarize(mean_estimator = mean(estimate)) %>% 
  ggplot(aes(x = truth, y = mean_estimator, fill = truth)) +
  geom_col() + 
  xlab("True Value of Beta_1") +
  ylab("Average estimate of Beta_1") +
  scale_x_continuous(
    breaks = c(1, 2, 3, 4, 5, 6)
  )
```

